Learning CNN filters from user-drawn image markers for coconut-tree image classification

### 2020-07-23 IEEE-GRSL-Response-Letter.tex

\author{Italos Estilon de Souza and Alexandre Xavier Falc\~{a}o}

\date{Manuscript GRSL-00821-2020}

\begin{document}
\maketitle

Dear Editor, please find attached the revised version of our manuscript, GRSL-00821-2020. We thank the Editor and
reviewers for the opportunity and comments to improve the presentation of our work. We have addressed all raised issues and highlighted the changes in red. Our point-by-point response to the reviewers' comments follows next.

@@

\rwr{1} 
The manuscript entitled "Learning CNN ﬁlters from user-drawn image-makers for coconut-tree image classiﬁcation" presents a method that needs a minimal set of user-selected images to train the CNN’s feature extractor, reducing the number of required images to train the fully connected layers. The theme of this manuscript is interesting, and the text brings an important contribution to the remote sensing area using deep learning tools. However, to be considered adequated to publish in the journal "Geoscience and Remote Sensing Letters", it requires a revision on the text. The attached document presents all corrections and observations made for it. However, the inclusion of more experiments using more images to train the proposed model is a crucial issue in my point of view. This is important to prove the robustness of the proposed model.

Thanks for the comments. We have included more experiments to demonstrate the robustness of the proposed feature learning approach. The parts marked and commented by reviewer 1 in the manuscript are reproduced and addressed as follows.

@@

Thanks for the comments. We have included more experiments to demonstrate the robustness of the proposed feature learning approach. The parts marked and commented by reviewer 1 in the manuscript are reproduced and addressed as follows.

\eans

\rwr{1} A well-trained CNN model should detect image regions that best discriminate classes as positive activations, improve class separation by eliminating negative activations, and aggregate the resulting activations locally (pooling) and globally (flattening) into a highly dimensional and sparse feature space, in which a linear classifier should separate the classes with reasonable success.

\ans  Thanks. The sentence was long and unclear. We have simplified it and improved it in first paragraph of Section II.
\eans

@@

\rwr{1} with resolution under 10 $cm^2$; LAB color space; 11387 regions; MLP classifier;

\ans Thanks. These are minor corrections in different parts of the text that have been solved now. (Section III). \eans

### 2020-07-23 main.tex
Updates from Overleaf

The design of CNN models without a human as part of the training loop leaves several questions unanswered: (1) How to find a useful and simplified model for a given classification problem? (2) How to train that model from a minimum number of annotated images?  (3) Can the user explain the decisions of the model? (4) Can the model improve from label corrections? The first question requires human knowledge about CNN modeling and the problem of interest. The second one requires to reduce the human effort to train a CNN model. The third issue is related to human understanding. It may explore data visualization techniques to explain the model's decisions and to guide the user in the design of the model. The fourth question is also essential during training, and it is related to user control over the process. They all lead to the importance of involving human experts during the machine learning process. To mitigate such problems,

Explainable artificial intelligence (XAI) has appeared to address the problem and avoid wrong interpretations of the results~\cite{arrietaInfoFusion2020}. In methods of XAI, however, the importance of human participation during the model's training process has called little attention yet~\cite{AngelovNN2020}. A network designer must choose a suitable architecture and the training hyperparameters with limited control over the quality of the model's estimated weights.

In this paper, we present significant advances related to the above questions. First, we conceptually divide a CNN into convolutional layers for feature extraction and fully connected layers for feature space reduction and classification. Each convolutional layer contains a filter bank, an activation function, and alternative operations (e.g., pooling, batch normalization).  As the number of convolutional layers increases (deeper is the model), and higher is the number of annotated images required to train the model by backpropagation.

To reduce the need for large annotated training sets by exploiting the user knowledge, we present a method, called \textcolor{red}{\emph{Feature Learning from Image Markers} (FLIM)}, that needs a minimal set of images to learn the filters of each convolutional layer. The user selects the number of convolutional layers, their filter sizes, \textcolor{red}{a few images from a training set}, and draws markers in image regions that best discriminate the classes. \textcolor{red}{FLIM can find filters that enhance those regions at the output of the convolutional layers. It is based on clustering of patches extracted from marker pixels at the input of each convolutional layer.} The CNN architecture could also be optimized by observing its results on a small validation set or based on data visualization techniques~\cite{rauber2016visualizing}.

@@

We used a dataset from~\cite{8899005}, which contains regions of aerial images with and without coconut trees from the Kingdom of Tonga, as acquired in October 2017~\footnote{The images are available in \url{blog.werobotics.org/2018/01/11/open-ai-challenge-2}.}. Each region should be classified as containing one or none coconut tree, but part of coconut trees from adjacent regions might appears near the region's border. The Humanitarian OpenStreetMap community annotated the regions. The dataset consists of 13587 regions with $90 \times 90$ pixels \textcolor{red}{with spatial resolution $8$\si{\centi\metre}}, being 10268 and 3319 regions annotated as containing and not containing coconut trees, respectively. \textcolor{red}{The region images contain three channels: red, green, and blue. We converted them to the CIE L*a*b* color space and then normalized the channels within $[0,1]$}.

\textcolor{red}{
To assess the impact of FLIM in reducing the need for annotated samples, we
divided the experiments in three parts. First, we randomly selected 200 regions (100 from each class) for training, 1000 regions (500 from each class) for the choice of architecture, and left the remaining 12387 regions for testing. The architecture of the network was chosen based on empirical tests on the validation set. Second, we used the validation set to increase the size of the training set from 200 to 1200 images.

>>>>

The design of CNN models without a human as part of the training loop leaves several questions unanswered: (1) How to find a useful and simplified model for a given classification problem? (2) How to train that model from a minimum number of annotated images?  (3) Can the user explain the decisions of the model? (4) Can the model improve from label corrections? The first question requires human knowledge about CNN modeling and the problem of interest. The second one requires to reduce the human effort to train a CNN model. The third issue is related to human understanding. It may explore data visualization techniques to explain the model's decisions and to guide the user in the design of the model. The fourth question is also essential during training, and it is related to user control over the process. They all lead to the importance of involving human experts during the machine learning process. \textcolor{red}{Explainable artificial intelligence (XAI) has appeared to address such problems~\cite{rauber2016visualizing,arrietaInfoFusion2020}. However, the importance of human participation during the training process has called little attention~\cite{AngelovNN2020}. Exceptions rely on visual analytics techniques~\cite{GarciaIJCNN2019,SpinnerTVCG2020}, and we present here a first work in which the user has direct influence over the quality of the filters.}

First, we conceptually divide a CNN into convolutional layers for feature extraction and fully connected layers for feature space reduction and classification. Each convolutional layer contains a filter bank, an activation function, and alternative operations (e.g., pooling, batch normalization).  As the number of convolutional layers increases (deeper is the model), and higher is the number of annotated images required to train the model by backpropagation.

To reduce the need for large annotated training sets by exploiting the user knowledge, we present a method, called \textcolor{red}{\emph{Feature Learning from Image Markers} (FLIM)}, that needs a minimal set of images to learn the filters of each convolutional layer. The user selects the number of convolutional layers, their filter sizes, \textcolor{red}{a few images from a training set}, and draws markers in image regions that best discriminate the classes. \textcolor{red}{By clustering patches extracted from marker pixels at the input of each convolutional layer, FLIM finds filters that enhance those regions at the output of the convolutional layers.} The CNN architecture could also be optimized based on visual analytics~\cite{SpinnerTVCG2020}.

@@

We use a dataset from~\cite{8899005}, which contains regions of aerial images with and without coconut trees from the Kingdom of Tonga, as acquired in October 2017~\footnote{The images are available in \url{blog.werobotics.org/2018/01/11/open-ai-challenge-2}.}. Each region is classified as containing one or none coconut tree, but part of coconut trees from adjacent regions might appear near the region's border. The Humanitarian OpenStreetMap community annotated the regions. The dataset consists of 13587 regions with $90 \times 90$ pixels \textcolor{red}{with spatial resolution of $8$\si{\centi\metre}}, being 10268 and 3319 regions annotated as containing and not containing coconut trees, respectively. \textcolor{red}{The region images contain three channels: red, green, and blue. We convert them to the CIE L*a*b* color space and normalize each channel within $[0,1]$}.

\textcolor{red}{The experiments have been designed to demonstrate the following hinvolved a dataset with X coconut-tree images from two classes, coconut-tree (Y images) and non-coconut-tree (Z images). All experiments were repeated T times, using random splits into training and testing sets, for statistical analysis with p-value 0.05 in Wilcoxon signed-rank test. The results were found statistically significant. The experiments have been designed to assess the following hypothesis about FLIM:

It is possible to learn competitive image feature extractors from user-drawn markers in a few training images per class.

We selected only 4 images from a training set with 100 images per class for the user to draw markers. An image feature extractor was trained by FLIM and the training set was used to design SVM and MLP classifiers for comparison with a well-established CNN, VGG-16. VGG was trained from scratch and also pre-trained on the ImageNet dataset and fine-tuned on the 200 training images. The classification results on the test set with the remaining W images indicate that the image feature extractor produced by FLIM can provide higher f-score than the feature extractor of VGG, with and without fine-tuning, and for both classifiers, SVM and MLP. These results are shown in Table I when we compare FLIM+SVM x VGG-FE+SVM, FLIM+MLP x VGG, and FLIM+MLP x VGG-FT.

For a given network architecture, the image features obtained by FLIM can produce similar or better classification results than those obtained by backpropagation on a same network architecture.

The comparison in Table I between FLIM+MLP and FLIM-FT+MLP shows that fine-tuning cannot improve the results already obtained by FLIM+MLP. Similarly, the comparison between FLIM+MLP x CL+MLP in Table I shows that a network with the same architecture of FLIM+MLP, trained from scratch, cannot surpass its results.

The competitive classification results obtained by FLIM remain competitive independently of the number of samples used to train the classifier.

We then repeated the comparisons between FLIM+MLP, FLIM-FT+MLP, and VGG-FT, the most competitive approaches, for increasing training-set sizes from 200 to 1200, and FLIM+MLP remained superior. The f-score curves of the methods in Figure X confirms the quality of the image feature extractor obtained by FLIM.

By increasing the number of relevant images selected to draw markers, one can improve the image feature extractor produced by FLIM.

            We finally compared FLIM+SVM, for the sake of simplicity, with an increasing number of images for marker selection (x1, x2, x3 images). As shown in Table II for X training images used to train the SVM classifier, this hypothesis is true. It is important to note that the results are promising but the selection of relevant images for marker selection is still a topic that deserves further investigation.

To assess the impact of FLIM in reducing the need for annotated samples, we
divided the experiments in three parts. First, we randomly selected 200 regions (100 from each class) for training, 1000 regions (500 from each class) for the choice of architecture, and left the remaining 12387 regions for testing. The architecture of the network was chosen based on empirical tests on the validation set. Second, we used the validation set to increase the size of the training set from 200 to 1200 images.
